---
title: 'Impulsivity and Self-Control as Timeless Concepts: Simulations'
author: "MASKED"
date: '2022-07-05'
output: html_document
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
library(renv)
library(scales)
```

<<<<<<< HEAD


I provide two toy examples to illustrate points in the manuscript. The
parameter optimisation function works but is not ideal for scaling up. Please
contact me on if you would like a more robust version.

```{r Generate data}
k = .00545
delta = .003

data <-
  tibble(
    largerLater = rep(100, 6),
    delay = c(0, 7, 21, 60, 182.5, 365)
  )

data <- 
  data %>%
  mutate(hyperbolic_discountFactor = 1/(1+k*delay),
         exponential_discountFactor = exp(-delta*delay),
         Hyperbolic = largerLater*hyperbolic_discountFactor,
         Exponential = largerLater*exponential_discountFactor) %>%
  pivot_longer(cols = c(Hyperbolic, Exponential),
               names_to = "model",
               values_to = "discountedValue")

```


```{r Plot data}
# Extract colours from palette.
scales::viridis_pal(begin = .3, end = .7)(2)
cols <- 
  c("#35608DFF", "#43BF71FF")

# Create plot showcasing difference between two functions.

(modelComparisonPlot <- 
  ggplot(data, 
       aes(x = delay, y = discountedValue, col = model)) +
  geom_point(size = 2) +
  ylim(c(0, 100)) +
  theme_light() +
  labs(x = "Delay in Days",
       y = "Discounted Present Value",
       col = "Model") +
  geom_function(fun = function(x) 100/(1+k*x),
                aes(col = "Hyperbolic")) +
  geom_function(fun = function(x) 100*exp(-delta*x),
                aes(col = "Exponential")) +
  scale_color_viridis_d(begin = .3, end = .7) +
  theme(legend.position = "bottom")) +
  annotate(geom = "text", 
           x = 300, 
           y = 75, 
           label = "paste(italic(k)[hyp], \" = .00545 in \", days ^ -1)",
             #"paste(italic(k) = .00545, \"in days ^ -1\")",
           col = cols[2],
           size = 4,
           parse = TRUE) +
  annotate(geom = "text", 
           x = 300, 
           y = 82.5, 
           label = "paste(italic(k)[exp], \" = .003 in \", days ^ -1)",
           col = cols[1],
           size = 4,
           parse = T) +
  coord_fixed(2)

ggsave(here("figures/plot_model-comparison.png"), width = 6, height = 4)



```


```{r Generate data to be estimated by wrong model}
# I now make a plot to illustrate the point that the hyperbolic model parameter
# k is not able to capture convexity of the data without also increasing 
# steepness.

# Create less steep exponential function to generate comparison data

delta2 = .001

dataWrongModel <-
  tibble(
    largerLater = rep(100, 6),
    delay = c(0, 7, 21, 60, 182.5, 365)
  )

dataWrongModel <- 
  dataWrongModel %>%
  mutate(Shallow = largerLater*exp(-delta2*delay), 
         Steep = largerLater*exp(-delta*delay))%>%
  pivot_longer(cols = c(Shallow, Steep),
               names_to = "Discount Rate",
               values_to = "discountedValue") %>%

  mutate(ID = ifelse(`Discount Rate` == "Shallow", "P1", "P2")) %>%
  # Give fictional participants ids. 
  arrange(ID)

```





```{r Estimating hyperbolic discounting parameters}

# A function for estimating the parameter for the hyperbolic discounting model
least_squares_hyp <-
  function(params, data) {
    
    k <- params[1]
    
    ll <- data$largerLater
    t2 <- data$delay
    indifP <- data$discountedValue
    
    est_npv_ll <- 1/(1+(t2)*k)*ll # Hyp. Discounted value
    temporaryResults <<- 
      data %>% 
      cbind(est_npv_ll, rep(k, times = nrow(data))) 
    # can check out how it works with temporaryResults 
    return(sum((est_npv_ll-indifP)^2))
  }

resultP1 <- optim(c(.5),
                least_squares_hyp, 
                method = "L-BFGS-B", 
                lower = c(.0001), upper = c(.99), 
                control = c(maxit = 10000),
                data = dataWrongModel %>%
                  filter(ID == "P1"))

resultP2 <- optim(c(.5),
                least_squares_hyp, 
                method = "L-BFGS-B", 
                lower = c(.0001), upper = c(.99), 
                control = c(maxit = 10000),
                data = dataWrongModel %>%
                  filter(ID == "P2"))

dataWrongModel <- 
  dataWrongModel %>% 
  mutate(k = ifelse(ID == "P1", 
                    resultP1$par, 
                    resultP2$par))

```



```{r Plot data with wrong model}
# estimate discount rate for the wrong model

false_k1 = resultP1$par 
false_k2 = resultP2$par

# Create figure with the wrong model fitted to the data
(wrongModelPlot <- 
  ggplot(dataWrongModel %>%
           mutate(k = factor(round(k, 4)),
                  k_exp = factor(rep(c(.001, .003), each = 6))), 
       aes(x = delay, y = discountedValue, shape = k_exp)) +
  geom_point(size = 2, col = cols[1]) +
  ylim(c(0, 100)) +
  theme_light() +
  geom_function(fun = function(x) 100/(1+false_k1*x),
                col = cols[2],
                aes(linetype = "0.0013")
                ) +
  geom_function(fun = function(x) 100/(1+false_k2*x),
                col = cols[2],
                aes(linetype = "0.0044")
                ) +
  labs(x = "Delay in Days",
       y = "Discounted Present Value",
       col = "Discount Function",
       linetype = bquote(k[hyp]),
       shape = bquote(k[exp])) +
  theme(legend.position = "bottom")) +
  coord_fixed(2)

ggsave(here("figures/plot_wrong-model.png"), width = 6, height = 4)


```

### End of figures for paper

Here are some more models to play around with:

```{r hyperboloid plots}
kMG1 = .005
kMG2 = .1

s1 = .7
s2 = .01

# Myerson Green 1995 model
(hyperboloidPlot <- 
  ggplot(dataWrongModel %>%
           mutate(k = factor(round(k, 4)),
                  k_exp = factor(rep(c(.001, .003), each = 6))), 
       aes(x = delay, y = discountedValue, shape = k_exp)) +
  geom_point(size = 2, col = cols[1]) +
  ylim(c(0, 100)) +
  theme_light() +
  geom_function(fun = function(x) 100/(1+kMG1*x)^s2,
                col = cols[2],
                aes(linetype = "0.0013")
                ) +
  geom_function(fun = function(x) 100/(1+kMG2*x)^s2,
                col = cols[2],
                aes(linetype = "0.0044")
                ) +
  labs(x = "Delay in Days",
       y = "Discounted Present Value",
       col = "Discount Function",
       linetype = bquote(k[hyp]),
       shape = bquote(k[exp])) +
  theme(legend.position = "bottom"))

```


```{r Modified Rachlin model plot}
# Modified Rachlin plot

kMR1 = .001
kMR2 = .005
sMR1 = 1
sMR2 = .6


(plotModRach <- 
  ggplot(dataWrongModel %>%
           mutate(k = factor(round(k, 4)),
                  k_exp = factor(rep(c(.001, .003), each = 6))), 
       aes(x = delay, y = discountedValue, shape = k_exp)) +
  geom_point(size = 2, col = cols[1]) +
  ylim(c(0, 100)) +
  theme_light() +
  geom_function(fun = function(x) 100/(1+(kMR1*x)^sMR2),
                col = cols[2],
                aes(linetype = "0.0013")
                ) +
  geom_function(fun = function(x) 100/(1+(kMR2*x)^sMR2),
                col = cols[2],
                aes(linetype = "0.0044")
                ) +
  labs(x = "Delay in Days",
       y = "Discounted Present Value",
       col = "Discount Function",
       linetype = bquote(k[hyp]),
       shape = bquote(k[exp])) +
  theme(legend.position = "bottom"))

```

```{r constant sensitivity plot}

a1 = .003
a2 = .01
b1 = .7
b2 = 1
b3 = 1.3
b4 = .4

(plotConstantSensitivity <- 
  ggplot(dataWrongModel, 
       aes(x = delay, y = discountedValue)) +
  geom_point(size = 2) +
  ylim(c(0, 100)) +
  theme_light() +
  labs(x = "Delay in Days",
       y = "Discounted Present Value",
       col = "b") +
  geom_function(fun = function(x) 100*exp(-(a1*x)^b1),
                aes(col = ".7")) +
  geom_function(fun = function(x) 100*exp(-(a1*x)^b2),
                aes(col = "1")) +
  geom_function(fun = function(x) 100*exp(-(a1*x)^b3),
                aes(col = "1.3")) +
    geom_function(fun = function(x) 100*exp(-(a1*x)^b4),
                aes(col = ".4")) +
  scale_color_viridis_d(begin = .3, end = .7) +
  theme(legend.position = "bottom")) +
  annotate(geom = "text", 
           x = 300, 
           y = 75, 
           label = "paste(italic(a)[CS], \" = .003 in \", days ^ -1)",
             #"paste(italic(k) = .00545, \"in days ^ -1\")",
           col = "black",
           size = 4,
           parse = TRUE)


```



